Lesson 5
========================================================

### Multivariate Data
Notes:

```{r}
getwd()

list.files()

library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')

names(pf)
```

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
library(dplyr)

# ggplot(aes(x = gender, y = age),
#        data = subset(pf, !is.na(gender))) + geom_boxplot()

pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count  = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>%
  ungroup() %>% # to remove age layer
  arrange(age)

head(pf.fc_by_age_gender)


```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y = median_friend_count),
       data = pf.fc_by_age_gender)+
  geom_line(aes(color = gender))

```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
#install.packages('reshape2')
library(reshape2)

pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender,
                                  value.var = 'median_friend_count')

head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(aes(x = age, y = female/male),
       data = pf.fc_by_age_gender.wide) +
  geom_line() +
  geom_hline(yintercept = 1, alpha = 0.3, linetype = 2)

```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined  <- floor(2014 - pf$tenure/365)

head(pf)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
# Cut a Variable : is useful to make discreet variabl from continuose or numberical ones 
# summary(pf$year_joined)
# table(pf$year_joined)

pf$year_joined.bucket <- cut(pf$year_joined,
                             c(2004, 2009, 2011, 2012, 2014))

```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
#  year_joined.bucket : categorical variables

ggplot(aes(x = age, y = friend_count), 
              data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)

```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
#  year_joined.bucket : categorical variables

ggplot(aes(x = age, y = friend_count), 
              data = subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)+
  
  # add grand mean
  geom_line(stat = 'summary', fun.y = mean, linetype = 2)

```



***

### Friending Rate
Notes:

```{r Friending Rate}
# user with at least 1 day tenure
with ( subset(pf, pf$tenure >= 1), summary(friend_count / tenure))

```

***

### Friendships Initiated
Notes:

What is the median friend rate? 0.2205

What is the maximum friend rate? 417

```{r Friendships Initiated}

ggplot(aes(x = tenure, y = friendships_initiated/tenure), 
              data = subset(pf, tenure >= 1 ) ) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean)
  
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

```{r}
# load dataset
yo <- read.csv('yogurt.csv')
str(yo)

# change id from int to factor
yo$id <- factor(yo$id)
str(yo)

```
***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
#ead(yo)

ggplot(aes(x = price), data = yo) + 
geom_histogram(aes(fill = "red"), binwidth = 0.5) 
#   scale_x_continuous(breaks = 1:31)

```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
# summary(yo)
# length(unique(yo$price))
# table(yo$price)
# 
# str(yo)

yo <-  transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)

summary(yo$all.purchases)
```

***

### Prices over Time
Notes:

```{r Prices over Time}


ggplot(aes(x = all.purchases), data = yo) + 
geom_histogram(fill = I("Yellow"), binwidth = 1) 

```

```{r Prices over Time solution}


ggplot(aes(x = time, y = price), data = yo) + 
  geom_jitter(alpha = 1/4, shape = 21, fill = I('#F79420'))
#geom_histogram(fill = I("Yellow"), binwidth = 1) 

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}

# set seed for reproducible results

set.seed(4230)

sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(x = time, y = price), 
       data = subset(yo, id %in% sample.ids)) +
  facet_wrap(~ id)+
  geom_line() +
  geom_point(aes(size = all.purchases), pch = 1)

  #geom_jitter(alpha = 1/4, shape = 21, fill = I('#F79420'))



```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:




***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

