---
title: "P04_Explore_And_Summarize_Data"
author: "2016 Presidential Campaign Financial Contributions - All Candidates"
output: html_document
---
```{r setwd()}
```
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r setup}
library(ggplot2)
library(gridExtra) 
library(dplyr)
library(reshape2)
```


```{r Load the data}
fl = read.csv('P00000001-FL.csv', row.names = NULL)
colnames(fl) <- c(colnames(fl)[-1], "x")
fl$x <- NULL
```

## Dataset Overview
```{r}
dim(fl)
```

```{r column names}
names(fl)
```

```{r Contributions}
summary(fl)
```

Our dataset consists of 18 variables, with almost 426,057 observations.


Let us see some sample data
```{r 2 samples}
head(fl, 2)
```



According [FEC.gov](http://www.fec.gov/info/contriblimitschart1718.pdfm)  :
Individual's contribution limit is $2700 so I'll eliminate any negative amounts 
(refunds) and amounts that exceed the contribution's limit .

```{r clean data by filtering contb_receipt_amt btween 1 and 2700}
fl_above_zero <- filter(fl, fl$contb_receipt_amt > 0)
fl_within_FEC_limit <- filter(fl_above_zero, 
                              fl_above_zero$contb_receipt_amt <= 2700)

dim(fl_within_FEC_limit)

head(fl_within_FEC_limit)

```


## Univariate Plots Section
### Exploring recieved contribution amount 
```{r}
summary(fl_within_FEC_limit$contb_receipt_amt)
```
```{r}
 ggplot(aes(x = contb_receipt_amt), data = fl_within_FEC_limit) + 
   geom_histogram(fill = '#000000')
```


```{r}
 ggplot(aes(x = contb_receipt_amt), data = fl_within_FEC_limit) + 
   geom_histogram (fill = '#000000')+
  scale_x_continuous(limits = c(0, 2700), breaks = seq(0, 2700, 100))

```


 From above, We can see about 120,000 contributor donated about less $145 . 
 <br>
 <br>


```{r contb count for each cand_nm}
 ggplot(aes(x = cand_nm), data = fl_within_FEC_limit) + 
   geom_histogram (stat="count", fill = '#000000') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Hillary clinton recieved highest contribution number with almost 190,000 contribution.
Bernie sanders roughly reciedved 80,000 contribution, followed 
closely by Donland Trump with about 76,000 contribution.
 <br>
 <br>

Let's see how much in total each of candidates recieved overall :

```{r total contb amt for each candidates - histogram }
ggplot(aes(x  = cand_nm, y = contb_receipt_amt), data = fl_within_FEC_limit) +
  geom_bar( stat="summary", fun.y=sum, fill = '#000000') +
    ylab('total_contb_receipt_amt') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
Hillary Clinton earned the lion's share of recieved contribution amount among 
other candidates by recieving almost $22,500,000 M.
What surprised me here is that Jeb Bush who had few thousands contrintuons got way 
a lot of money comparing to other candidates who had better contribtuiobns numbers 
such as Bernie sanders and Marco Rubio.
 <br>
 <br>

Boxplot would give us a better visulization about how recived contribtuion amount distributed across candidates : 
```{r total contb amt for each candidates - boxplot}
ggplot(aes(x  = cand_nm, y = contb_receipt_amt), data = fl_within_FEC_limit) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

We have a lot of outliers, Our boxplo needs some tweaks to make it more readable :
```{r contributions distrbution for candidates }
ggplot(aes(x  = cand_nm, y = contb_receipt_amt), data = fl_within_FEC_limit) +
  scale_y_log10() +
  geom_boxplot(alpha = 0.2, color = 'Black') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
 <br>
 <br>

Let's the average contribution amount for each candidates :
```{r mean contb amt for each candidates - histogram }
ggplot(aes(x  = cand_nm, y = contb_receipt_amt), data = fl_within_FEC_limit) +
  geom_bar( stat="summary", fun.y=mean, fill = '#000000') +
  ylab('contb_receipt_amt_mean') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
 <br>
 <br>
 <br>
 <br>

### Now, I'm going to explore top 10 cities with most contribution number/amount :
```{r 10 top cities by contribution count}
count_by_city <- fl_within_FEC_limit %>%
  group_by(contbr_city) %>%
  count() %>%
  arrange(desc(n)) %>%
  top_n(10)
head(count_by_city)
```
```{r}
ggplot(aes(x  = contbr_city, y= n), data = count_by_city) +
  geom_col(fill = '#000000')
```

Looking at our barchart, I was surprised that Jacksonville wasn't the city with 
highest contribution number since it's the largest city in Florida.
Seeing Miami, Tampa and Orlando among the top 5 made a lot of sense. 
However, I was a bit shocked that Naples slightly surpassed Jacksonville. 
One justification I could think of is that Naples is considered of the richest 
cities in Florida state. 
 <br>
 <br>
 <br>

```{r 10 top cities by sum contribution amount}
sum_by_city <- fl_within_FEC_limit %>%
  group_by(contbr_city) %>%
  summarise(sum  = sum(contb_receipt_amt), 
            n = n()) %>%
  arrange(desc(sum)) %>%
  top_n(10)
```



```{r}
ggplot(aes(x  = contbr_city, y= sum), data = sum_by_city) +
  geom_col(fill = '#000000') 
```
Interestingly, two small cities Naples and Boca Raton had significant 
contribution comparing to other large cities such as Orlando and Jacksonville. 

<br>
<br>


## Bivariate Plots Section

### Exploring contribtuin amount and contributin date :

Let's see how contribution is distrubted across years
```{r convert contb_receipt_dt to date }
fl_within_FEC_limit$contb_receipt_dt <- as.Date(fl_within_FEC_limit$contb_receipt_dt, 
                                                "%d-%b-%y")
```

```{r add new column : year}
fl_within_FEC_limit$Year <- format(as.Date(fl_within_FEC_limit$contb_receipt_dt), 
                                   "%Y")
```

```{r let us plot group by year}
ggplot(aes(x = Year, y = contb_receipt_amt), data = fl_within_FEC_limit) + 
   geom_boxplot()
```

I expected 2016 to have highest contribution amount mean comparing to other years, 
since it's the election year and presidential race gets closer as we approach 
nominees announcement date. However, data proved me wrong as I clearly see 2014 
has the highest contribtuon amount mean .
 <br>
 <br>

Let me explore election year (2016) further :
```{r plot contb_receipt_amt of 2016}
ggplot(aes(x = contb_receipt_dt, y = contb_receipt_amt), 
       data = subset(fl_within_FEC_limit, 
                     fl_within_FEC_limit$Year == '2016')) + 
   geom_point( alpha=0.2,fill = '#000000') +
    geom_smooth()

```




```{r add new column : Month_year}
fl_within_FEC_limit$Month_year <- format(as.Date(fl_within_FEC_limit$contb_receipt_dt), 
                                         "%b %Y")
```

```{r group contribution amount by Month_year}
fl_within_FEC_limit.amt_by_Month_year <- subset(fl_within_FEC_limit, 
                                                fl_within_FEC_limit$Year 
                                                == '2016') %>%
  group_by(Month_year) %>%
  summarise(contb_receipt_amt_sum = sum(contb_receipt_amt),
            contb_receipt_amt_mean = mean(contb_receipt_amt),
            contb_receipt_amt_median = median(contb_receipt_amt),
            n = n()) %>%
  arrange(Month_year)

```


 <br>
 <br>
 Let's see monthly sum/mean/median of contribution amount for 2016 :
```{r plot Month_year for 2016}
p1 <- ggplot(aes(x = Month_year, y = contb_receipt_amt_sum, group = 1), 
             data = fl_within_FEC_limit.amt_by_Month_year) + 
  geom_line()

p2 <- ggplot(aes(x = Month_year, y = contb_receipt_amt_mean, group = 1), 
             data = fl_within_FEC_limit.amt_by_Month_year) + 
   geom_line()

p3 <- ggplot(aes(x = Month_year, y = contb_receipt_amt_median, group = 1), 
             data = fl_within_FEC_limit.amt_by_Month_year) + 
   geom_line()

grid.arrange(p1, p2, p3, ncol=1)

```

<br>
<br>
<br>

### Exploring contribtuin amount and contbr_occupation :

```{r}
sum_amt_by_occupation <- fl_within_FEC_limit %>%
  filter(contbr_occupation != 'INFORMATION REQUESTED',
         contbr_occupation != 'INFORMATION REQUESTED PER BEST EFFORTS') %>%
  group_by(contbr_occupation) %>%
  summarise(contb_receipt_amt_sum  = sum(contb_receipt_amt),
            contb_receipt_amt_mean = mean(contb_receipt_amt),
            n = n()) %>%
  arrange(desc(contb_receipt_amt_sum)) %>%
  top_n(10)
```

```{r plot total amt by occupation using bar chart}
ggplot(aes(x=contbr_occupation, y=contb_receipt_amt_sum), 
       data=sum_amt_by_occupation) +
    geom_bar(stat='identity', fill = '#000000')  
```

Retireds poured a lot of moeny into the elections, marking themselves as the 
biggest contributors among differebt occuptaions. Way behind, comes attorneys 
and then homemakers in the third place.

<br>
<br>

Let us what was the mean contribtuions among those top occupations :
```{r plot mean amt by occupation using bar chart}
ggplot(aes(x=contbr_occupation, y=contb_receipt_amt_mean), 
       data=sum_amt_by_occupation) +
    geom_bar(stat='identity', fill = '#000000')  
```

Homemakers were the dark horse here by contributing about $460 on avrage. 
Attorneys followed with average contribution less than $400


<br>
<br>

## Multivariate Plots Section
Let's look into top 10 occupations and thier impact on two final nominees, Donald Trump and Hillary Clinton

```{r top 10 occupations - for Donald Trump and Hillary Clinton}
group_by_multi <- fl_within_FEC_limit %>%
  group_by(cand_nm, contbr_occupation ) %>%
   filter(cand_nm %in% c("Trump, Donald J.", "Clinton, Hillary Rodham"),
          contbr_occupation %in% sum_amt_by_occupation$contbr_occupation) %>%
  summarise(contb_receipt_amt_sum  = sum(contb_receipt_amt),
            contb_receipt_amt_mean = mean(contb_receipt_amt),
            n = n()) %>%
  arrange(desc(contb_receipt_amt_sum))  %>%
  top_n(10)
```


```{r plot sum amt of top 10 occupations - for Donald Trump and Hillary Clinton}
ggplot(aes(x = contbr_occupation, y = contb_receipt_amt_sum),
       data = group_by_multi)+
  geom_bar(stat='identity',position = 'dodge', aes(fill = cand_nm)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Retireds were the most highest contributor to both nominees. Attorneys played 
a heavy role contribtuting way much money to Hillary comapring to fractions Donlad 
recieved from them.
<br>
<br>

```{r plot mean amt of top 10 occupations - for Donald Trump and Hillary Clinton}
ggplot(aes(x = contbr_occupation, y = contb_receipt_amt_mean),
       data = group_by_multi)+
  geom_line(aes(color = cand_nm, group = cand_nm)) +
  #geom_bar(stat='identity',position = 'dodge', aes(fill = cand_nm))+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Trump has better average recived contribtuion amount than Hillary across all top 10 occupation
except Attornys.
Attronys donated about $400 on average to Hillary, and about $250 on average to Trump.


```{r top 10 cities - Donald Trump and Hillary Clinton}
group_by_city_can <- fl_within_FEC_limit %>%
  group_by(cand_nm, contbr_city) %>%
   filter(cand_nm %in% c("Trump, Donald J.", "Clinton, Hillary Rodham"),
          contbr_city %in% count_by_city$contbr_city) %>%
  summarise(contb_receipt_amt_sum  = sum(contb_receipt_amt),
            contb_receipt_amt_mean = mean(contb_receipt_amt),
            n = n()) %>%
  arrange(desc(contb_receipt_amt_sum))  %>%
  top_n(10)

head(group_by_city_can)
```

<br>
<br>

Let's see how both nominees did in top 10 contributing cities : 

```{r plot sum amt of top 10 cities -  Donald Trump and Hillary Clinton}
ggplot(aes(x = contbr_city, y = contb_receipt_amt_sum),
       data = group_by_city_can)+
  geom_bar(stat='identity' , aes(fill = cand_nm)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
Hillary led in Miami with about $22,000,000 M recicved contribtuion, 
and Trump recvivced less than $5000,000 M. 
However, Naples played in favor of Trump where he collected about $7000,000 .

<br>
<br>
<br>
<br>

## Final Plots and Summary

### Plot One

```{r}
ggplot(aes(x  = contbr_city, y= sum), data = sum_by_city) +
  geom_col(fill = '#099DD9') +
  scale_y_continuous(breaks = seq(0, 5000000, 500000)) +
  xlab('City of contribution') +
  ylab('Total contribution amount in USD')
```
Rich citites in Florida played major role contributing to candidates. 
We can see how rich, small cities liek Naples and Boca Raton have more i
mpact than major cities such as Jacksonville.

<br>
<br>

### Plot Two

```{r plot contb_receipt_amt of 2016 - jitter}
ggplot(aes(x = contb_receipt_dt, y = contb_receipt_amt), 
       data = subset(fl_within_FEC_limit, 
                     fl_within_FEC_limit$Year == '2016')) + 
  scale_y_log10() +
geom_point(alpha=0.3, position='jitter') +
  geom_smooth()
```
Contributin amounts started going up somewhere aroud Septmeber, 
reaching the peak in October and then started decreasing toward election month, Nov.

<br>
<br>

### Plot Three

```{r }
ggplot(aes(x = contbr_occupation, y = contb_receipt_amt_sum),
       data = group_by_multi)+
  geom_bar(stat='identity', position = 'dodge', aes(fill = cand_nm)) +
  scale_fill_manual(values=c("Blue","red")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  xlab('Occupations') +
  ylab('Contribution recieved amount in USD')+
  labs(fill="Candidate name")
  
```
Hillary topped Trump in all top 10 occuptins.

<br>
<br>
<br>
<br>

## Reflection


Florida state 2016 election contribution dataset contains more than 417573 items 
which were captured between 2013-10-01 and 2016-12-31. 
To begin, I omitted contribution above FEC limit which is $2700, and any refunes.

The average contribuion was around $145, and the minimum recieved contribution was $0.15.


Most of the contribution was made to hillary clinton, 
bernie sanders and Donland Trump respectivly.

Contributors from Miami, Tampa and Naples contributed the major portion of recived money.


On average, contribution amount was about $300 in 2014 which was best avergae contributon comparing to other years.

Retireds, attorneys and homemakers were major players by contriuting most of the moeny recived. 
On average, homemakers donated about 
$460 and attoreys 
about $390 .


It'd be very interesting to look how condidates performed in other swing states. Further, explore contribution based on the candidate's party and his/her background.


Dealing with US politics was a new experience for me. I had to read and understand how election contribution system works, what the limitations are, how candidates target supports, etc. 
For example, I had to understand the limitations on allowed maximum contribution amount, and why I have negative numbers in the dataset and how to deal with them.







