---
title: "P04_Explore_And_Summarize_Data"
author: "2016 Presidential Campaign Financial Contributions - All Candidates"
output: html_document
---
```{r setwd()}
```
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r setup}
library(ggplot2)
library(gridExtra) 
library(dplyr)
library(reshape2)
```


```{r Load the data}
fl = read.csv('P00000001-FL.csv', row.names = NULL)
colnames(fl) <- c(colnames(fl)[-1], "x")
fl$x <- NULL
```

## Dataset Overview
```{r}
dim(fl)
```

```{r column names}
names(fl)
```

```{r Contributions}
summary(fl)
```

Our dataset consists of 18 variables, with almost 426,057 observations.


Let us see some sample data
```{r 2 samples}
head(fl, 2)
```



According [FEC.gov](http://www.fec.gov/info/contriblimitschart1718.pdfm)  :
Individual's contribution limit is $2700 so I'll eliminate any negative amounts 
(refunds) and amounts that exceed the contribution's limit .

```{r clean data by filtering contb_receipt_amt btween 1 and 2700}
fl_above_zero <- filter(fl, fl$contb_receipt_amt > 0)
fl_within_FEC_limit <- filter(fl_above_zero, 
                              fl_above_zero$contb_receipt_amt <= 2700)

dim(fl_within_FEC_limit)

head(fl_within_FEC_limit)

```


## Univariate Plots Section
### Exploring recieved contribution amount 
```{r}
summary(fl_within_FEC_limit$contb_receipt_amt)
```
```{r}
 ggplot(aes(x = contb_receipt_amt), data = fl_within_FEC_limit) + 
   geom_histogram(fill = '#000000')
```


```{r}
 ggplot(aes(x = contb_receipt_amt), data = fl_within_FEC_limit) + 
   geom_histogram (fill = '#000000')+
  scale_x_continuous(limits = c(0, 2700), breaks = seq(0, 2700, 100))

```


 Frm above, We can see about 120,000 contributor donated about less $145 . 
 <br>
 <br>


```{r contb count for each cand_nm}
 ggplot(aes(x = cand_nm), data = fl_within_FEC_limit) + 
   geom_histogram (stat="count", fill = '#000000') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


```{r contb candidates }
ggplot(aes(x  = cand_nm, y = contb_receipt_amt), data = fl_within_FEC_limit) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r contributions distrbution for candidates }
ggplot(aes(x  = cand_nm, y = contb_receipt_amt), data = fl_within_FEC_limit) +
  scale_y_log10() +
  geom_boxplot(alpha = 0.2, color = 'Black') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Hillary Clinton earned the lion's share of recieved contribution amount among 
other candidates, followed by Bernard Sanders and Donald Trump. 
 <br>
 <br>
 <br>
 <br>

### Now, I'm going to explore top 10 cities with most contribution number/amount :
```{r 10 top cities by contribution count}
count_by_city <- fl_within_FEC_limit %>%
  group_by(contbr_city) %>%
  count() %>%
  arrange(desc(n)) %>%
  top_n(10)
```
```{r}
ggplot(aes(x  = contbr_city, y= n), data = count_by_city) +
  geom_col(fill = '#000000')
```

Looking at our barchart, I was surprised that Jacksonville wasn't the city with 
highest contribution number since it's the largest city in Florida.
Seeing Miami, Tampa and Orlando among the top 5 made a lot of sense. 
However, I was a bit shocked that Naples slightly surpassed Jacksonville. 
One justification I could think of is that Naples is considered of the richest 
cities in Florida state. 
 <br>
 <br>
 <br>

```{r 10 top cities by sum contribution amount}
sum_by_city <- fl_within_FEC_limit %>%
  group_by(contbr_city) %>%
  summarise(sum  = sum(contb_receipt_amt), 
            n = n()) %>%
  arrange(desc(sum)) %>%
  top_n(10)
```



```{r}
ggplot(aes(x  = contbr_city, y= sum), data = sum_by_city) +
  geom_col(fill = '#000000') 
```
Interestingly, two small cities Naples and Boca Raton had significant 
contribution comparing to other large cities such as Orlando and Jacksonville. 

<br>
<br>


## Bivariate Plots Section

### Exploring contribtuin amount and contributin date :

Let's see how contribution is distrubted across years
```{r convert contb_receipt_dt to date }
fl_within_FEC_limit$contb_receipt_dt <- as.Date(fl_within_FEC_limit$contb_receipt_dt, 
                                                "%d-%b-%y")
```

```{r add new column : year}
fl_within_FEC_limit$Year <- format(as.Date(fl_within_FEC_limit$contb_receipt_dt), 
                                   "%Y")
```

```{r let us plot group by year}
ggplot(aes(x = Year, y = contb_receipt_amt), data = fl_within_FEC_limit) + 
   geom_boxplot()
```

I expected 2016 to have highest contribution amount mean comparing to other years, 
since it's the election year and presidential race gets closer as we approach 
nominees announcement date. However, data proved me wrong as I clearly see 2014 
has the highest contribtuon amount mean .
 <br>
 <br>

Let me explore election year (2016) further :
```{r plot contb_receipt_amt of 2016}
ggplot(aes(x = contb_receipt_dt, y = contb_receipt_amt), 
       data = subset(fl_within_FEC_limit, 
                     fl_within_FEC_limit$Year == '2016')) + 
   geom_point( alpha=0.2,fill = '#000000') +
    geom_smooth()

```




```{r add new column : Month_year}
fl_within_FEC_limit$Month_year <- format(as.Date(fl_within_FEC_limit$contb_receipt_dt), 
                                         "%b %Y")
```

```{r group contribution amount by Month_year}
fl_within_FEC_limit.amt_by_Month_year <- subset(fl_within_FEC_limit, 
                                                fl_within_FEC_limit$Year 
                                                == '2016') %>%
  group_by(Month_year) %>%
  summarise(contb_receipt_amt_sum = sum(contb_receipt_amt),
            contb_receipt_amt_mean = mean(contb_receipt_amt),
            contb_receipt_amt_median = median(contb_receipt_amt),
            n = n()) %>%
  arrange(Month_year)

```


 <br>
 <br>
 Let's see monthly sum/mean/median of contribution amount for 2016 :
```{r plot Month_year for 2016}
p1 <- ggplot(aes(x = Month_year, y = contb_receipt_amt_sum, group = 1), 
             data = fl_within_FEC_limit.amt_by_Month_year) + 
  geom_line()

p2 <- ggplot(aes(x = Month_year, y = contb_receipt_amt_mean, group = 1), 
             data = fl_within_FEC_limit.amt_by_Month_year) + 
   geom_line()

p3 <- ggplot(aes(x = Month_year, y = contb_receipt_amt_median, group = 1), 
             data = fl_within_FEC_limit.amt_by_Month_year) + 
   geom_line()

grid.arrange(p1, p2, p3, ncol=1)

```

<br>
<br>
<br>

### Exploring contribtuin amount and contbr_occupation :

```{r}
sum_amt_by_occupation <- fl_within_FEC_limit %>%
  filter(contbr_occupation != 'INFORMATION REQUESTED',
         contbr_occupation != 'INFORMATION REQUESTED PER BEST EFFORTS') %>%
  group_by(contbr_occupation) %>%
  summarise(contb_receipt_amt_sum  = sum(contb_receipt_amt),
            contb_receipt_amt_mean = mean(contb_receipt_amt),
            n = n()) %>%
  arrange(desc(contb_receipt_amt_sum)) %>%
  top_n(10)
head(sum_amt_by_occupation, 15)
```

```{r plot total amt by occupation using bar chart}
ggplot(aes(x=contbr_occupation, y=contb_receipt_amt_sum), 
       data=sum_amt_by_occupation) +
    geom_bar(stat='identity', fill = '#000000')  
```

Retireds poured a lot of moeny into the elections, marking themselves as the 
biggest contributors among differebt occuptaions. Way behind, comes attorneys 
and then homemakers in the third place.

<br>
<br>
Let us what was the mean contribtuions among those top occupations :
```{r plot mean amt by occupation using bar chart}
ggplot(aes(x=contbr_occupation, y=contb_receipt_amt_mean), 
       data=sum_amt_by_occupation) +
    geom_bar(stat='identity', fill = '#000000')  
```

Homemakers were the dark horse here by contributing about $460 on avrage. 
Attorneys followed with average contribution less than $400


<br>
<br>

## Multivariate Plots Section

```{r}
group_by_multi <- fl_within_FEC_limit %>%
  group_by(cand_nm, contbr_occupation ) %>%
   filter(cand_nm %in% c("Trump, Donald J.", "Clinton, Hillary Rodham"),
          contbr_occupation %in% sum_amt_by_occupation$contbr_occupation) %>%
  summarise(contb_receipt_amt_sum  = sum(contb_receipt_amt),
            contb_receipt_amt_mean = mean(contb_receipt_amt),
            n = n()) %>%
  arrange(desc(contb_receipt_amt_sum))  %>%
  top_n(10)

head(group_by_multi, 100)
dim(group_by_multi)
```

```{r }
ggplot(aes(x = contbr_occupation, y = contb_receipt_amt_sum),
       data = group_by_multi)+
  geom_bar(stat='identity',position = 'dodge', aes(fill = cand_nm)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

```{r }
ggplot(aes(x = contbr_occupation, y = contb_receipt_amt_mean),
       data = group_by_multi)+
  geom_line(aes(color = cand_nm, group = cand_nm)) +
  #geom_bar(stat='identity',position = 'dodge', aes(fill = cand_nm))+
    theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Looks like Donlad Trump was so popular among retired contributor. 
On the other hand, Sandaers got a lot of contribution from unemployed folks.








## Final Plots and Summary

### Plot One

```{r}
ggplot(aes(x  = contbr_city, y= sum), data = sum_by_city) +
  geom_col(fill = '#099DD9') +
  scale_y_continuous(breaks = seq(0, 5000000, 500000)) +
  xlab('City of contribution') +
  ylab('Total contribution amount in USD')
```
Rich citites in Florida played major role contributing to candidates. 
We can see how rich, small cities liek Naples and Boca Raton have more i
mpact than major cities such as Jacksonville.


### Plot Two
```{r plot contb_receipt_amt of 2016 - jitter}
ggplot(aes(x = contb_receipt_dt, y = contb_receipt_amt), 
       data = subset(fl_within_FEC_limit, 
                     fl_within_FEC_limit$Year == '2016')) + 
geom_point(alpha=0.3, position='jitter') +
  geom_smooth()
```
Contributin amounts started going up somewhere aroud Septmeber, 
reaching the peak in October and then started decreasing toward election month, Nov.




### Plot Three
```{r }
ggplot(aes(x = contbr_occupation, y = contb_receipt_amt_sum),
       data = group_by_multi)+
  geom_bar(stat='identity', position = 'dodge', aes(fill = cand_nm)) +
  xlab('Contributor Occupation') +
  ylab('Contribution recieved amount in USD')
  
```


## Reflection


Florida state 2016 election contribution dataset contains more than 417573 items 
which were captured between 2013-10-01 and 2016-12-31. 
To begin, I omitted contribution above FEC limit which is $2700, and any refunes.

The average contribuion was around $145, and the minimum recieved contribution was $0.15.


Most of the contribution was made to hillary clinton, 
bernie sanders and Donland Trump respectivly.

Contributors from Miami, Tampa and Naples contributed the major portion of recived money.


On average, contribution amount was about $300 in 2014 which was best avergae contributon comparing to other years.

Retireds, attorneys and homemakers were major players by contriuting most of the moeny recived. On average, homemakers donated about $460 and attoreys about $390 .

Candidate Benjamin carson recived the lowest contribution overall in Florida, and most of the contribution he recieved was from retireds.


It'd be very interesting to look how condidates performed in other swing states. Further, explore contribution based on the candidate's party and his/her background.





