Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:

```{r}
getwd()

list.files()

library(ggplot2)
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')

names(pf)
```
***

### Scatterplots
Notes:
- examine relationship btw 2 continouse variables
```{r Scatterplots}
qplot(x = age, y = friend_count, data = pf)

# alternative (ggplot1)


```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
summary(pf$age)

ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point() + # aes() is wrapper
  xlim(13,90)

```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20) + # aes() is wrapper
  xlim(13,90)

ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_jitter(color = 'red', alpha = 1/20) + # aes() is wrapper
  xlim(13,90)

```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(aes(x = age, y = friend_count), data = pf) + 
  geom_point(alpha = 1/20) +
  xlim(13,90) +
  coord_trans(y = "sqrt") 

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}

```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
# ggplot(aes(x = age, y = friendships_initiated), data = pf) + 
#   geom_point(alpha = 1/20, position = 'jitter')  +
#   coord_trans(y = 'sqrt')
## above code will produce error cuz we have casues where y is zero !

ggplot(aes(x = age, y = friendships_initiated), data = pf) + 
  geom_point(alpha = 1/20, position = position_jitter(h = 0))  +
  coord_trans(y = 'sqrt')
```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
#install.packages('dplyr')
library(dplyr)

pf.fc_by_age <- pf %>%
  group_by(age) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age)

head(pf.fc_by_age, 20)
```

Create your plot!

```{r Conditional Means Plot}
ggplot(aes(x = age, y = friend_count_mean ), data = pf.fc_by_age) +
  geom_line() #geom_point()
```

***

### Overlaying Summaries with Raw Data
Notes:


```{r Overlaying Summaries with Raw Data}
ggplot(aes(x = age, y = friendships_initiated), data = pf) + 
    coord_cartesian(xlim = c(13, 90), ylim = c(0, 1000)) +
    geom_point(alpha = 0.05, 
               position = position_jitter(h = 0),
               color = 'Orange')  +
  geom_line(stat = 'summary' , fun.y = mean)+
    geom_line(stat = 'summary' , fun.y = quantile, fun.args = list(probs = 0.1), linetype = 2, color = 'blue') +
  geom_line(stat = 'summary' , fun.y = quantile, fun.args = list(probs = 0.5), linetype = 2, color = 'blue') +
  geom_line(stat = 'summary' , fun.y = quantile, fun.args = list(probs = 0.9), linetype = 2, color = 'blue')
```





#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}
cor.test(pf$age, pf$friend_count, method = 'pearson')

#alternative
with(pf, cor.test(age, friend_count, method = 'pearson'))
```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(  subset(pf, age <= 70) , cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:
Pearson product-memont correlation measures the strength of relationship between any 2 variables

***

## Create Scatterplots
Notes:

```{r}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + 
    geom_point()

```

***

### Strong Correlations
Notes:
The correlation coefficient is invariant under a linear transformation of either X or Y, and the slope of the regression line when both X and Y have been transformed to z-scores is the correlation coefficient. 


```{r Strong Correlations}
ggplot(aes(x = www_likes_received, y = likes_received), data = pf) + 
  geom_point() +
  xlim(0, quantile(pf$www_likes_received, 0.95)) +
  ylim(0, quantile(pf$www_likes_received, 0.95)) +
  geom_smooth(method = 'lm', color = 'red')

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
# correlation coefficent
cor.test(pf$www_likes_received, pf$likes_received)

  

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:
correlation could help us to decide which variables are related

```{r More Caution With Correlation}
#install.packages('alr3')
library(alr3)

data(Mitchell)
?Mitchell
```

Create your plot!

```{r Temp vs Month}
#head(Mitchell, 50)

ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)
#0.057 very weak correlation

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
#range(Mitchell$Month)
ggplot(aes(x = Month, y = Temp), data = Mitchell) +
  geom_point() +
  scale_x_continuous(breaks = seq(0, 203, 12)) # to make month discreet (every 12 months) from 0 to 203, break every 12

```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}
# add new column
pf$age_with_months <- pf$age + (1 - pf$dob_month / 12) 


```

***

### Age with Months Means

```{r Age with Months Means}
pf.fc_by_age_months <- pf %>%
  group_by(age_with_months) %>%
  summarise(friend_count_mean = mean(friend_count),
            friend_count_median = median(friend_count),
            n = n()) %>%
  arrange(age_with_months)

head(pf.fc_by_age_months)

```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}
ggplot(aes(x = age_with_months, y = friend_count_mean ), 
       data = subset(pf.fc_by_age_months, age_with_months < 71 )) +
  geom_line() #geom_point()

```
54
***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}
library(gridExtra)
p1 <- ggplot(aes(x = age, y = friend_count_mean ), 
             data = subset(pf.fc_by_age, age < 70)) +
  geom_line() +
  geom_smooth()

p2 <- ggplot(aes(x = age_with_months, y = friend_count_mean ), 
       data = subset(pf.fc_by_age_months, age_with_months < 71 )) +
  geom_line() +
  geom_smooth()


grid.arrange(p2, p1, ncol = 1)


```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:
- scatter plots
- conditional summaries (liek : means)
- correlation coefficient


***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

